### 2021-03-04 paper3/paper3.tex
Merge references
Fix headings and remove duplicate bib entries

%\icmltitle{Learning with Perfect Bags:\\
%Addressing Hidden Stratification with Zero Labelled Data}
\textsc{Authors}:\\
Thomas Kehrenberg$^1$, Myles Bartlett$^1$, Oliver Thomas$^1$, and Novi Quadrianto$^1$ \\
\textsc{Affiliations}:\\
$^1$ Predictive Analytics Lab (PAL), University of Sussex, Brighton, UK\\
\textsc{Conference}:\;\;\textit{European Conference on Computer Vision} (ECCV), 2020 \\
% \textsc{DOI}:\;\;10.3389/frai.2020.00033
\textsc{Note}:\;\;The appendix has been included as section~\ref{sec:sec:zsf-appendix}.

@@

\section{Introduction}%
\label{sec:introduction}
Machine learning has been deployed in safety-critical applications such as medicine (e.g. \citet{DunYiLanReetal19}), and socially important contexts such as the allocation of healthcare, education, and credit (\eg\ \citet{HurAde17,RagBarKleLev20}). 
%
Efficiency can be improved, costs can be reduced, and personalization of services and products can be greatly enhanced -- these are some of the drivers for the widespread development and deployment of machine learning algorithms. 
%

Algorithms such as classifiers, however, are trained from large amount of labelled data, and are typically trained to optimise \emph{global} metrics such as average classification accuracy.
%
% \begin{figure*}[t]
% \centering
% \includegraphics[width=0.6\textwidth,page=2]{figures/ideal.pdf}
% \caption{Overview of the zero-shot stratification problem. Training dataset (pairs of input data $x$ and class label $y$) will only contain data points that are labeled ``yes'' by the decision policy. This systematic bias might result in a subgroup(s) to have zero labeled data (in the example above, ``married'' subgroup has zero labeled data). The subgrouping information $s$ is unavailable/unlabeled at deployment time, and is only partially labeled at train time.}
% \vspace{-0.5cm}
% \label{fig:censoring}
% \end{figure*}
In many real-world classification tasks, each labelled class consists of multiple semantically distinct subclasses, or subgroups.
For example, the ``dog'' class label can have finer-grained intra-class variations, such as ``dog indoor'' and ``dog outdoor''.
%
This finer-grained subgrouping information is typically unavailable/unlabeled (e.g. \citet{SohDunAngGuetal20,NamChaAhnLeeetal20}).
%
The standard training process brings about two inter-connected challenges: a) classifiers often under-perform on important hidden subgroups (\emph{hidden stratification}) \cite{RayDunCarRe20,SohDunAngGuetal20};
and b) \emph{systematic bias}~\cite{KalZho18} affects whether or not entire collections of data points appear in the training dataset,
and can make the classifier unprepared for treating those subgroups in the eventual deployment setting
% and can hamper attempts to correct for bias in the eventual deployment data
(\emph{residual bias}). %\cite{KalZho18}.

We are interested in hidden stratification in which, for some of the subgroups, labelled training data is only available with a certain outcome, or labelled training samples are not available at all, due to systematic bias.
This can be seen as a strong sampling bias.
%
For instance, data on loan defaults can only be collected on those loan
applicants who were approved in the past \cite{KalZho18}.
%
Here a loan decision policy specifies whether an individual will be included in the training dataset.
%
Individuals can be thought of as belonging to a specific subgroup such as ``married'' or ``not married'' and systematic bias produced by a historical decision policy may result in the ``not married'' subgroup having poor, or altogether non-existent, representation.
%For instance, a medical imaging model trained to classify between ``benign'' and ``abnormal'' lesions may achieve high overall performance, yet consistently mislabel a rare but critical abnormal subgroup as ``benign'' \cite{DunYiLanReetal19}.
%
%Each class actually composes of ``no-patch'' and ``patch'' hidden subgroups, and the ``no-patch'' subgroup is further splitted into “histopathology” and “non-histopathology”.
%
%A past decision policy of \emph{not} requiring histopathology (biopsy \& pathologist referral) specifies whether an individual will be included in the training dataset.
%
%This systematic bias might lead to the no-patch subgroup to have thin or non-existent labeled data.
%
We formalise this stratification problem as a data setting where a decision policy can lead to one or more subgroups having no labelled data.

@@

We aim to convert our unlabelled data into a collection of \emph{perfect bags} \citep{KleMulRag16,chouldechova17}, \ie\ sample sets in which the class label $y$ and subgroup label $s$ are independent (\ie\ $y\perp s$).
Making use of the terminology from \emph{multiple-instance learning}, our batches comprise a certain number of bags which are a collection of samples.
%
We will then use these perfect bags as the inductive bias for learning the disentangled representations. 
%
How can we construct these perfect bags in the absence of any labelled data?
%
We assume that the number of subgroups is known \emph{a priori}\footnote{Relaxing this assumption represents a clear avenue for future work. We elaborate this in the limitation and intended use sec.~\ref{sec:limitations}.}. 
%We elaborate this further in our experiments and discussions sections.}.
%corresponding to the diverse demographic groups in the real-world population in which our machine learning system will be deployed. 
%
We then apply unsupervised k-means clustering, or a \emph{semi}-supervised clustering based on rank statistics; the latter allows incorporating annotations from the training data when forming the clusters.
%
Once the clusters have been found, we can sample from each cluster at an equal rate to form balanced (\ie\,\emph{perfect}) bags and use them as input for learning a disentangled representation. 
%
%See fig. \ref{fig:teaser} for an overview of our learning with invisible demographic framework.
We cast the problem of disentangling as one of distribution matching and propose an adversarial learning approach.
%a parametric approach based on adversarial learning. 
In the standard GAN setting, the discriminator assigns a score to each sample corresponding to the perceived probability that it was drawn from the true data distribution, and not the generator's.
In contrast, we train a discriminator to assign scores at the level of bags of samples, with a bag being deemed authentic if it was drawn from the originally unbiased distribution and not the de-biased one (with the de-biaser playing the role of the generator). To do so, we take inspiration from set-classification and multiple-instance learning and equip the discriminator with a learnable attention mechanism to model interdependencies between samples in a bag.

@@

\textbf{Related work.}
We describe related work in two areas: zero-shot learning and semi-supervised learning.
%and disentangled representation learning.

\textbf{On zero-shot learning.}
The setting with incomplete training data, where we aim to account for seen and unseen outcomes is also known as \emph{generalised zero-shot learning}. 
Traditionally, zero-shot learning transfers knowledge from classes for which we have training data to classes for which we do not, via auxiliary knowledge, e.g. via prototype examples \citep{larochelle2008}, intermediate class description such as semantic attributes \citep{lampert2009, xian2018zero}, word2vec embeddings \citep{bucher2019}. 
Our method similarly uses a collection of perfect bags as a source of auxiliary knowledge but in contrast to generalised zero-shot learning, our perfect bag is an unlabelled pool of data, where class descriptions are unknown.

\textbf{On semi-supervised learning.} 
\citet{WicPanTri19} proposed a semi-supervised method that can successfully harness unlabelled data to correct for the selection bias and label bias in the training data.
%
The unlabelled data, despite not containing the class label $y$, \emph{is} labelled in terms of the subgroup label $s$. 
%
Our setting is significantly harder because there is no label information about $y$ and $s$ in the perfect bag.

\textbf{On disentangled representations learning.}
\citet{LocAbbRaiBauetal19} suggested that disentanglement in representation learning may be a useful property to remove algorithmic bias when subgroup information is not observed.
%
In order for disentangled representations to reduce algorithmic bias without the knowledge of subgroup label $s$, they have to assume that the class label $y$ and the subgroup label $s$ are independent, i.e. $y\perp s$.
%
Though, in many real-world tasks, the variable $s$ is correlated with the variable $y$, and therefore unsupervised methods are not suitable  \citep{JaiWuAbdNat18,JaiWuAbdNat19}. 
%
Indeed, experiments in \citet{LocAbbRaiBauetal19} were wholly done with procedurally generated synthetic datasets involving 2D and 3D shapes. 
%
Without some supervision or inductive bias, disentangled representation methods would not solve the issue of algorithmic fairness with invisible demographics \citep{LacBauLucGeletal19}.

%\begin{wrapfigure}{!t}{0.6\textwidth}
\begin{figure}[ht]
\centering
    %\includegraphics[width=0.9\textwidth]{figures/SSL-framework}
    % \includegraphics[width=\textwidth]{figures/SSL-framework-withPred.pdf}
    \includegraphics[width=\textwidth]{figures/zsf_diagram.pdf}
    \caption{%
    The main components involved in our proposed disentangling method, $f_y$ (debiaser) and $h$ (discriminator). The debiaser is trained to produce encodings, $z_y$ of the data that are invariant to the source dataset and thereby the subgroups identifying it. In order to determine whether a bag of encodings originates from the training set or the deployment set, the discriminator performs an attention-weighted aggregation over the bag dimension to model interdependencies between the samples. In the case of Coloured MNIST where {\color{purple}purple} fours constitute the missing subgroup, the discriminator can identify an encoding of a bag from the training set by the absence of such samples so long as colour information is detectable in $z_y$, serving as an error signal for the debiaser.
    }%
    \label{fig:architecture}
\end{figure}%
% \end{figure}
%\end{wrapfigure}
% \hfill
% \begin{figure}
\begin{figure}[ht]
    \centering
    % if you want to make changes to the diagram, it is located here:
    % https://drive.google.com/file/d/1M37m4rVj1EGrFJb9QzK32bs8swepiaZ1/view?usp=sharing
    \includegraphics[width=\textwidth]{figures/fdm.pdf}
    \caption{%
    Illustration of distribution matching.
    In this example, the training set is lacking {\color{purple}purple} 4's.
    By enforcing the subspace $z_y$ to have the same distribution for both the training and deployment set, the model is encouraged to learn a representation that is invariant to colour (or $s$ in general.)
    For this to work, it is crucial that the bags be approximately balanced (perfect bags).
    }%
    \label{fig:matching-diagram}
% \end{subfigure}
% \caption{%
% Overview of the disentangling framework with a perfect (balanced) bag as an inductive bias.
% }%
% \label{fig:framework}
\end{figure}

\section{Methodology}\label{sec:methodology}
\subsection{Theoretical background}
In this section, we first formalise the problem of hidden stratification with zero data (zero-label stratification) and the related issue of algorithmic bias. 
%
We then theoretically motivate the idea of perfect bags for reducing algorithmic bias, and their use as an inductive bias for disentanglement.

\textbf{Zero-label stratification and algorithmic bias.}
Let $S$ denote a set of discrete-valued subgroup labels of the associated domains $\gS$.
%
% $S$ can take the values taken by a single subclass label, or, $S = S_1 \times S_2 \times \ldots \times S_p$ with $S_1, \ldots, S_p$ be discrete-valued subclass labels more generally.
%
$X$, with the associated domain $\gX$, represents other attributes of the data. % that may or may not be influenced by subgroup labels $S$.
%
Let $\gY$ denote the space of class labels for a classification task; $\gY = \{0,1\}$ for binary classification or $\gY = \{1,2,\ldots,C_{\text{cls}}\}$ for multi-class classification.
%
For ease of exposition, we assume that we have multiple sources $\Omega$ of samples, one for each combination of class-label $y$ and subgroup-label $s$.
That is, we have:
\begin{align}
\Omega_{y=y',s=s'},\quad\quad\forall y'\in\mathcal{Y},\forall s'\in\mathcal{S},
\end{align}
where, for example, the source $\Omega_{y=0,s=0}$ supplies all data points with class label $y=0$ and subgroup label $s=0$. 
%
As in a standard supervised learning task, we have access to a labelled training set $\gD_{tr} =\{(x_i, s_i, y_i)\}$, that is used to learn a model $M:\gX \rightarrow \gY$. 
$\gD_{tr}$ is composed of several sources, but lacks samples from some of the sources:
\begin{align}
\exists y'\in \gY,\exists s'\in \gS: \gD_{tr}\cap \Omega_{y=y',s=s'} = \varnothing.
\end{align}
For example, we might be missing samples from two sources: $\Omega_{y=0,s=0}$ and $\Omega_{y=1,s=0}$. 
%
In binary classification, this corresponds to no labelled data for the subgroup label $s=0$, a setting we refer to as \emph{missing subgroup} (MS).
%
Other times, we may observe a one-sided (negative) outcome for the subgroup label $s=0$
(\ie, we have $\gD_\mathit{tr} \cap\Omega_{y=1,s=0} = \varnothing$), giving rise to a setting we refer to as \emph{subgroup bias} (SB).

@@

%Semi-supervised learning \cite{ChaSchZie06} can alleviate the issue of unfairness to the \emph{invisibles} by mixing labeled with unlabeled data, which is usually much cheaper to obtain. 
%
We propose to alleviate the issue of bias against missing subgroups by mixing labelled data with unlabelled data that is usually much cheaper to obtain \cite{ChaSchZie06}. 
%
In this paper, we refer to this set of \emph{unlabelled} data as the
deployment set\footnote{In our experiments, we report accuracy and bias metrics on another independent test set instead of on the unlabelled data that is available at training time.} $\gD_{dep} =\{(x_i)\}$.
%_{i=1}^{N^c}
%
This 
%context 
deployment set has overlap with all sources:
\begin{align}
  \gD_{dep} \cap \Omega_{y=y',s=s'} \neq \varnothing\quad\quad\forall y'\in\gY,\forall s'\in\gS~.
\end{align}
%The deployment set is much like the deployment set: 
Importantly, the deployment set has no information about class labels $y$ or the subgroup labels $s$.

\textbf{Relation to algorithmic fairness.}
Training set bias also affects what has been termed \emph{algorithmic fairness},
which is commonly expressed in terms of the predicted class $\hat{y}$ of a machine learning model $M$.
We adopt a statistical notion of algorithmic fairness in which outcomes are balanced under certain conditions between groups of data points with different subgroup labels.
% We adopt a statistical notion of algorithmic fairness in which it balances a certain condition between groups of data points with different subgroup labels.
%
% The variable $\hat{y}$ below is the prediction of a machine learning model $M$.
%
Several statistical bias measures have been proposed
% equality of acceptance rate (demographic parity), equality of true positive rates, equality of true negative rate, equality of positive predicted value, equality of negative predicted value
\citep{kamiran2012data,HarPriSre16,ZafValRodGum17b,chouldechova17,RagBarKleLev20}
(shown below for the case where $s$ and $y$ are binary):
\begin{align}
& P(\hat{y}=1|s=0)=P(\hat{y}=1|s=1)\label{eq:AR}\\
& P(\hat{y}=1|s=0,y)=P(\hat{y}=1|s=1,y) \label{eq:TPR} \\
%& P(\hat{y}=0|s=0,y=0)=P(\hat{y}=0|s=1,y=0) %& \text{(equality of true negative rate)} \label{eq:TNR}\\
& P(y=1|s=0,\hat{y})=P(y=1|s=1,\hat{y}) \label{eq:PPV}
%& P(y=0|s=0,\hat{y}=0)=P(y=0|s=1,\hat{y}=0) & \text{(equality of negative predicted value)}\label{eq:NPV}.
\end{align}
(\ref{eq:AR}) is equality of positive rate; (\ref{eq:TPR}) is equality of true positive/negative rate; (\ref{eq:PPV}) is equality of positive/negative predicted value.
Generally, these statistical notions can be expressed in terms of different (conditional) independence statements between the involved random variables \citep{BarHarNar19}:
% $\hat{y}\perp s$ (equality of acceptance rate), $\hat{y}\perp s\ |\ y$ (equality of true positive and true negative rate), and $y\perp s\ |\ \hat{y}$ (equality of positive and negative predicted value).
$\hat{y}\perp s$ (\eqref{eq:AR}), $\hat{y}\perp s\ |\ y$ (\eqref{eq:TPR}), and $y\perp s\ |\ \hat{y}$ (\eqref{eq:PPV}).
If our training set has no positive outcome for the subgroup label $s=0$, i.e. $\Omega_{y=1,s=0} = \varnothing$, the true positive rate for this subgroup will suffer, and therefore we will likely not be able to satisfy, among others, equality of true positive rate.
In the experimental section, we use metrics based on these equalities to quantify how strongly the predictions are affected by the dataset bias.

>>>>

%\icmltitle{Learning with Perfect Bags:\\
%Addressing Hidden Stratification with Zero Labelled Data}
\textsc{Authors}:\\
Thomas Kehrenberg$^1$, Viktoriia Sharmanska$^1$, Myles Bartlett$^1$, and Novi Quadrianto$^1$ \\
\textsc{Affiliations}:\\
$^1$ Predictive Analytics Lab (PAL), University of Sussex, Brighton, UK\\
\textsc{Conference}:\;\; Submitted to \textit{Internation Conference on Machine Learning} (ICML), 2021 \\
% \textsc{DOI}:\;\;10.3389/frai.2020.00033
\textsc{Note}:\;\;The appendix has been included as section~\ref{sec:zsf-appendix}.

@@

\section{Introduction}%
\label{sec:introduction}
Machine learning has been deployed in safety-critical applications such as medicine (e.g. \citep{DunYiLanReetal19}), and socially important contexts such as the allocation of healthcare, education, and credit (\eg\ \citep{HurAde17,RagBarKleLev20}). 
%
Efficiency can be improved, costs can be reduced, and personalization of services and products can be greatly enhanced -- these are some of the drivers for the widespread development and deployment of machine learning algorithms. 
%

Algorithms such as classifiers, however, are trained from large amount of labelled data, and are typically trained to optimise \emph{global} metrics such as average classification accuracy.
%
% \begin{figure*}[t]
% \centering
% \includegraphics[width=0.6\textwidth,page=2]{figures/ideal.pdf}
% \caption{Overview of the zero-shot stratification problem. Training dataset (pairs of input data $x$ and class label $y$) will only contain data points that are labeled ``yes'' by the decision policy. This systematic bias might result in a subgroup(s) to have zero labeled data (in the example above, ``married'' subgroup has zero labeled data). The subgrouping information $s$ is unavailable/unlabeled at deployment time, and is only partially labeled at train time.}
% \vspace{-0.5cm}
% \label{fig:censoring}
% \end{figure*}
In many real-world classification tasks, each labelled class consists of multiple semantically distinct subclasses, or subgroups.
For example, the ``dog'' class label can have finer-grained intra-class variations, such as ``dog indoor'' and ``dog outdoor''.
%
This finer-grained subgrouping information is typically unavailable/unlabeled (e.g. \citep{SohDunAngGuetal20,nam2020learning}).
%
The standard training process brings about two inter-connected challenges: a) classifiers often under-perform on important hidden subgroups (\emph{hidden stratification}) \citep{RayDunCarRe20,SohDunAngGuetal20};
and b) \emph{systematic bias}~\citep{kallus2018residual} affects whether or not entire collections of data points appear in the training dataset,
and can make the classifier unprepared for treating those subgroups in the eventual deployment setting
% and can hamper attempts to correct for bias in the eventual deployment data
(\emph{residual bias}). %\cite{kallus2018residual}.

We are interested in hidden stratification in which, for some of the subgroups, labelled training data is only available with a certain outcome, or labelled training samples are not available at all, due to systematic bias.
This can be seen as a strong sampling bias.
%
For instance, data on loan defaults can only be collected on those loan
applicants who were approved in the past \citep{kallus2018residual}.
%
Here a loan decision policy specifies whether an individual will be included in the training dataset.
%
Individuals can be thought of as belonging to a specific subgroup such as ``married'' or ``not married'' and systematic bias produced by a historical decision policy may result in the ``not married'' subgroup having poor, or altogether non-existent, representation.
%For instance, a medical imaging model trained to classify between ``benign'' and ``abnormal'' lesions may achieve high overall performance, yet consistently mislabel a rare but critical abnormal subgroup as ``benign'' \cite{DunYiLanReetal19}.
%
%Each class actually composes of ``no-patch'' and ``patch'' hidden subgroups, and the ``no-patch'' subgroup is further splitted into “histopathology” and “non-histopathology”.
%
%A past decision policy of \emph{not} requiring histopathology (biopsy \& pathologist referral) specifies whether an individual will be included in the training dataset.
%
%This systematic bias might lead to the no-patch subgroup to have thin or non-existent labeled data.
%
We formalise this stratification problem as a data setting where a decision policy can lead to one or more subgroups having no labelled data.

@@

We aim to convert our unlabelled data into a collection of \emph{perfect bags} \citep{kleinberg2016inherent,chouldechova2017fair}, \ie\ sample sets in which the class label $y$ and subgroup label $s$ are independent (\ie\ $y\perp s$).
Making use of the terminology from \emph{multiple-instance learning}, our batches comprise a certain number of bags which are a collection of samples.
%
We will then use these perfect bags as the inductive bias for learning the disentangled representations. 
%
How can we construct these perfect bags in the absence of any labelled data?
%
We assume that the number of subgroups is known \emph{a priori}\footnote{Relaxing this assumption represents a clear avenue for future work. We elaborate this in the limitation and intended use sec.~\ref{sec:limitations}.}. 
%We elaborate this further in our experiments and discussions sections.}.
%corresponding to the diverse demographic groups in the real-world population in which our machine learning system will be deployed. 
%
We then apply unsupervised k-means clustering, or a \emph{semi}-supervised clustering based on rank statistics; the latter allows incorporating annotations from the training data when forming the clusters.
%
Once the clusters have been found, we can sample from each cluster at an equal rate to form balanced (\ie\,\emph{perfect}) bags and use them as input for learning a disentangled representation. 
%
%See fig. \ref{fig:teaser} for an overview of our learning with invisible demographic framework.
We cast the problem of disentangling as one of distribution matching and propose an adversarial learning approach.
%a parametric approach based on adversarial learning. 
In the standard GAN setting, the discriminator assigns a score to each sample corresponding to the perceived probability that it was drawn from the true data distribution, and not the generator's.
In contrast, we train a discriminator to assign scores at the level of bags of samples, with a bag being deemed authentic if it was drawn from the originally unbiased distribution and not the de-biased one (with the de-biaser playing the role of the generator). To do so, we take inspiration from set-classification and multiple-instance learning and equip the discriminator with a learnable attention mechanism to model interdependencies between samples in a bag.

@@

\section{Related work.}
We describe related work in two areas: zero-shot learning and semi-supervised learning.
%and disentangled representation learning.

\subsection{On zero-shot learning.}
The setting with incomplete training data, where we aim to account for seen and unseen outcomes is also known as \emph{generalised zero-shot learning}. 
Traditionally, zero-shot learning transfers knowledge from classes for which we have training data to classes for which we do not, via auxiliary knowledge, e.g. via prototype examples \citep{larochelle2008}, intermediate class description such as semantic attributes \citep{lampert2009, xian2018zero}, word2vec embeddings \citep{bucher2019}. 
Our method similarly uses a collection of perfect bags as a source of auxiliary knowledge but in contrast to generalised zero-shot learning, our perfect bag is an unlabelled pool of data, where class descriptions are unknown.

\subsection{On semi-supervised learning.} 
\citet{wick2019unlocking} proposed a semi-supervised method that can successfully harness unlabelled data to correct for the selection bias and label bias in the training data.
%
The unlabelled data, despite not containing the class label $y$, \emph{is} labelled in terms of the subgroup label $s$. 
%
Our setting is significantly harder because there is no label information about $y$ and $s$ in the perfect bag.

\subsection{On disentangled representations learning.}
\citet{LocAbbRaiBauetal19} suggested that disentanglement in representation learning may be a useful property to remove algorithmic bias when subgroup information is not observed.
%
In order for disentangled representations to reduce algorithmic bias without the knowledge of subgroup label $s$, they have to assume that the class label $y$ and the subgroup label $s$ are independent, i.e. $y\perp s$.
%
Though, in many real-world tasks, the variable $s$ is correlated with the variable $y$, and therefore unsupervised methods are not suitable  \citep{jaiswal2018unsupervised,JaiWuAbdNat19}. 
%
Indeed, experiments in \citet{LocAbbRaiBauetal19} were wholly done with procedurally generated synthetic datasets involving 2D and 3D shapes. 
%
Without some supervision or inductive bias, disentangled representation methods would not solve the issue of algorithmic fairness with invisible demographics \citep{LacBauLucGeletal19}.

%\begin{wrapfigure}{!t}{0.6\textwidth}
\begin{figure}[t]
\centering
    %\includegraphics[width=0.9\textwidth]{figures/SSL-framework}
    % \includegraphics[width=\textwidth]{figures/SSL-framework-withPred.pdf}
    \includegraphics[width=\textwidth]{figures/zsf_diagram.pdf}
    \caption{%
    The main components involved in our proposed disentangling method, $f_y$ (debiaser) and $h$ (discriminator). The debiaser is trained to produce encodings, $z_y$ of the data that are invariant to the source dataset and thereby the subgroups identifying it. In order to determine whether a bag of encodings originates from the training set or the deployment set, the discriminator performs an attention-weighted aggregation over the bag dimension to model interdependencies between the samples. In the case of Coloured MNIST where {\color{purple}purple} fours constitute the missing subgroup, the discriminator can identify an encoding of a bag from the training set by the absence of such samples so long as colour information is detectable in $z_y$, serving as an error signal for the debiaser.
    }%
    \label{fig:architecture}
\end{figure}%
% \end{figure}
%\end{wrapfigure}
% \hfill
% \begin{figure}
\begin{figure}[t]
    \centering
    % if you want to make changes to the diagram, it is located here:
    % https://drive.google.com/file/d/1M37m4rVj1EGrFJb9QzK32bs8swepiaZ1/view?usp=sharing
    \includegraphics[width=0.5\textwidth]{figures/fdm.pdf}
    \caption{%
    Illustration of distribution matching.
    In this example, the training set is lacking {\color{purple}purple} 4's.
    By enforcing the subspace $z_y$ to have the same distribution for both the training and deployment set, the model is encouraged to learn a representation that is invariant to colour (or $s$ in general.)
    For this to work, it is crucial that the bags be approximately balanced (perfect bags).
    }%
    \label{fig:matching-diagram}
% \end{subfigure}
% \caption{%
% Overview of the disentangling framework with a perfect (balanced) bag as an inductive bias.
% }%
% \label{fig:framework}
\end{figure}

\section{Methodology}\label{sec:methodology}
\subsection{Theoretical background}
In this section, we first formalise the problem of hidden stratification with zero data (zero-label stratification) and the related issue of algorithmic bias. 
%
We then theoretically motivate the idea of perfect bags for reducing algorithmic bias, and their use as an inductive bias for disentanglement.\\

\noindent\textsc{Zero-label stratification and algorithmic bias.}
% \textsc{Stratification.}
\;\; Let $S$ denote a set of discrete-valued subgroup labels of the associated domains $\gS$.
%
% $S$ can take the values taken by a single subclass label, or, $S = S_1 \times S_2 \times \ldots \times S_p$ with $S_1, \ldots, S_p$ be discrete-valued subclass labels more generally.
%
$X$, with the associated domain $\gX$, represents other attributes of the data. % that may or may not be influenced by subgroup labels $S$.
%
Let $\gY$ denote the space of class labels for a classification task; $\gY = \{0,1\}$ for binary classification or $\gY = \{1,2,\ldots,C_{\text{cls}}\}$ for multi-class classification.
%
For ease of exposition, we assume that we have multiple sources $\Omega$ of samples, one for each combination of class-label $y$ and subgroup-label $s$.
That is, we have:
\begin{align}
\Omega_{y=y',s=s'},\quad\quad\forall y'\in\mathcal{Y},\forall s'\in\mathcal{S},
\end{align}
where, for example, the source $\Omega_{y=0,s=0}$ supplies all data points with class label $y=0$ and subgroup label $s=0$. 
%
As in a standard supervised learning task, we have access to a labelled training set $\gD_{tr} =\{(x_i, s_i, y_i)\}$, that is used to learn a model $M:\gX \rightarrow \gY$. 
$\gD_{tr}$ is composed of several sources, but lacks samples from some of the sources:
\begin{align}
\exists y'\in \gY,\exists s'\in \gS: \gD_{tr}\cap \Omega_{y=y',s=s'} = \varnothing.
\end{align}
For example, we might be missing samples from two sources: $\Omega_{y=0,s=0}$ and $\Omega_{y=1,s=0}$. 
%
In binary classification, this corresponds to no labelled data for the subgroup label $s=0$, a setting we refer to as \emph{missing subgroup} (MS).
%
Other times, we may observe a one-sided (negative) outcome for the subgroup label $s=0$
(\ie, we have $\gD_\mathit{tr} \cap\Omega_{y=1,s=0} = \varnothing$), giving rise to a setting we refer to as \emph{subgroup bias} (SB).

@@

%Semi-supervised learning \cite{ChaSchZie06} can alleviate the issue of unfairness to the \emph{invisibles} by mixing labeled with unlabeled data, which is usually much cheaper to obtain. 
%
We propose to alleviate the issue of bias against missing subgroups by mixing labelled data with unlabelled data that is usually much cheaper to obtain \citep{ChaSchZie06}. 
%
In this paper, we refer to this set of \emph{unlabelled} data as the
deployment set\footnote{In our experiments, we report accuracy and bias metrics on another independent test set instead of on the unlabelled data that is available at training time.} $\gD_{dep} =\{(x_i)\}$.
%_{i=1}^{N^c}
%
This 
%context 
deployment set has overlap with all sources:
\begin{align}
  \gD_{dep} \cap \Omega_{y=y',s=s'} \neq \varnothing\quad\quad\forall y'\in\gY,\forall s'\in\gS~.
\end{align}
%The deployment set is much like the deployment set: 
Importantly, the deployment set has no information about class labels $y$ or the subgroup labels $s$.

\paragraph{Relation to algorithmic fairness.}
Training set bias also affects what has been termed \emph{algorithmic fairness},
which is commonly expressed in terms of the predicted class $\hat{y}$ of a machine learning model $M$.
We adopt a statistical notion of algorithmic fairness in which outcomes are balanced under certain conditions between groups of data points with different subgroup labels.
% We adopt a statistical notion of algorithmic fairness in which it balances a certain condition between groups of data points with different subgroup labels.
%
% The variable $\hat{y}$ below is the prediction of a machine learning model $M$.
%
Several statistical bias measures have been proposed
% equality of acceptance rate (demographic parity), equality of true positive rates, equality of true negative rate, equality of positive predicted value, equality of negative predicted value
\citep{kamiran2012data,hardt2016equality,zafar2017fairnesstreatment,chouldechova2017fair,RagBarKleLev20}
(shown below for the case where $s$ and $y$ are binary):
\begin{align}
& P(\hat{y}=1|s=0)=P(\hat{y}=1|s=1)\label{eq:AR}\\
& P(\hat{y}=1|s=0,y)=P(\hat{y}=1|s=1,y) \label{eq:TPR} \\
%& P(\hat{y}=0|s=0,y=0)=P(\hat{y}=0|s=1,y=0) %& \text{(equality of true negative rate)} \label{eq:TNR}\\
& P(y=1|s=0,\hat{y})=P(y=1|s=1,\hat{y}) \label{eq:PPV}
%& P(y=0|s=0,\hat{y}=0)=P(y=0|s=1,\hat{y}=0) & \text{(equality of negative predicted value)}\label{eq:NPV}.
\end{align}
(\ref{eq:AR}) is equality of positive rate; (\ref{eq:TPR}) is equality of true positive/negative rate; (\ref{eq:PPV}) is equality of positive/negative predicted value.
Generally, these statistical notions can be expressed in terms of different (conditional) independence statements between the involved random variables \citep{BarHarNar19}:
% $\hat{y}\perp s$ (equality of acceptance rate), $\hat{y}\perp s\ |\ y$ (equality of true positive and true negative rate), and $y\perp s\ |\ \hat{y}$ (equality of positive and negative predicted value).
$\hat{y}\perp s$ (\eqref{eq:AR}), $\hat{y}\perp s\ |\ y$ (\eqref{eq:TPR}), and $y\perp s\ |\ \hat{y}$ (\eqref{eq:PPV}).
If our training set has no positive outcome for the subgroup label $s=0$, i.e. $\Omega_{y=1,s=0} = \varnothing$, the true positive rate for this subgroup will suffer, and therefore we will likely not be able to satisfy, among others, equality of true positive rate.
In the experimental section, we use metrics based on these equalities to quantify how strongly the predictions are affected by the dataset bias.

\paragraph{Perfect bag.}
We call a sampled set for which $y\perp s$ holds, a perfect bag \citep{chouldechova2017fair,kleinberg2016inherent}.
Such sets are also very desirable when training algorithmically fair classifiers
(see the \emph{sampling} method in \citet{kamiran2012data}).
%
%If we have access to perfect bags, we could equalise true positive/negative rates (\eqref{eq:TPR}) and also equalise positive/negative predicted values (\eqref{eq:PPV}) for all demographic groups.
%This can be shown by using the sum and product rule of conditional probabilities, e.g. \citet{KanRotZia19}.
%We consider a binary-valued subgroup label, $s\sp{\prime}$ versus $s\sp{\prime\prime}$. 
%For $s\sp{\prime}$, we can compute: 
%$P(y=1|\hat{y}=1,s\sp{\prime})=$
%%\begin{align}
%$\nicefrac{P(\hat{y}=1|y=1,s\sp{\prime})P(y=1|s\sp{\prime})}{\left(A + B\right)}$, with $A = P(\hat{y}=1|y=1,s\sp{\prime})P(y=1|s\sp{\prime})$ and $B = P(\hat{y}=1|y=0,s\sp{\prime})(1-P(y=1|s\sp{\prime}))$.
%%\end{align}
%We can do similarly for $s\sp{\prime\prime}$.
%The conditional probability on the left hand side is 
%a positive predicted value, and this quantity can be expressed in terms of true positive/negative rates and the base (prior) rate, shown on the right hand side. 
%If we have a perfect bag ($y\perp s$ holds, which means equal base rates $P(y=1|s\sp{\prime}) = P(y=1|s\sp{\prime\prime})$), an equality in the true positive/negative rates will give us an equality in the positive/negative predicted values.
%Similarly, with a perfect bag, we can equalise true positive/negative rates (eq. \ref{eq:TPR}) and also positive rates (eq. \ref{eq:AR}) for all demographic groups.
%From the sum probability rule, we have:
%$
%P(\hat{y}=1|s\sp{\prime})
%= P(\hat{y}=1|y=1,s\sp{\prime})P(y=1|s\sp{\prime}) + P(\hat{y}=1|y=0,s\sp{\prime})(1-P(y=1|s\sp{\prime}))
%$ for $s\sp{\prime}$ value, and accordingly for $s\sp{\prime\prime}$ value.
%Here, a positive rate on the left hand side is related to true positive/negative rates and the base (prior) rate as shown on the right hand side. 
%
Ideally, we would like to sample our deployment dataset as perfect bags.
However, the deployment set is unlabelled and is unlikely to be perfect in practice. 
Instead, we pursue learning under zero-label systematic bias as learning disentangled representations with a collection of \emph{approximately} perfect bags. We show that the disentangling procedure is robust enough to work with this relaxation, but that performance scales with how well the deployment set is balanced.

\paragraph{Disentangled representation.}
Disentanglement-learning aims to find a factorised representation of a data point $x$ through mapping functions $f_i$ such that $f_i(x)=z_i$ where $z_1,z_2,\ldots,z_p$ are $p$ distinct (independent) factors of variations, which together form $x$.
%
We can formalise this intuitive definition using group and representation theories \citep{HigAmoPfaRacetal18}, or using structural causal models \citep{SutMilSchBau19}.
%
Specifically for this paper, we would like to split the data representation into two factors as $f_y(x)=z_y$ and $f_s(x)=z_s$ where $z_y$ contains factors that are relevant for $y$-prediction and $z_s$ contains factors related to the subgroup label $s$.
%
Since $s$ is correlated with the class label $y$, we need annotations of the undesired nuisance variable $s$ \citep{jaiswal2018unsupervised,JaiWuAbdNat19} to be successful in using disentanglement learning methods for zero-label stratification. %\citet{jaiswal2018unsupervised,JaiWuAbdNat19}.
%
We have some annotations of subgroup label $s$ in the training set $\gD_{tr} =\{(x_i, s_i, y_i)\}$, however, crucially, due to systematic bias, this set is missing certain subgroups.
%
We have all subgroups in the deployment set $\gD_{dep} =\{(x_i)\}$, 
%(also in the deployment set $\Dcal_{tr} =\{(x_i)\}$), 
though, the challenge is that the subgrouping information is unavailable or hidden at the deployment time.
%
In the following section, we show that we can still leverage the 
deployment set for learning the disentangled representations.
